---
title: "Regional Well Being"
author: Akash Mittal, Maged Saeed Abdo Mostafa Kharshom, Precious Prince, Franco Reinaldo
  Bonifacini
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(openxlsx) # to read and write Excel files (.xlsx) in R
library(tidyr) # data format conversion, tidying and transforming data
library(tidyverse) # Whole package of libraries (dplyr, tidyr, ggplot2, readr, purrr, tibble, stringr, forcats)

library(ggplot2) # for plotting
library(reshape2) # for reshaping and transforming data
library(fpp3) # used for forecasting in R
library(GGally) # functionality of ggplot2 as it provides additional geoms and functions for data visualization
library(ggcorrplot) # used for the heatmap correlation matrix
library(fmsb) # used for creating radar charts
library(plotly) # used for the interactive plot
library(randomcoloR) # used for generating colors

library(lme4) # for linear mixed effects modeling + functions for fitting and analyzing linear mixed effects models
library(insight) # functions for model-agnostic interpretation and visualization of regression models
```

# Abstract

|   This project aims to **estimate the effect of various factors such as health, education, and income among others, on the life satisfaction of people in different regions from different countries**. For this, we carried out a robust approach with MEMs & Robust Estimators. Also, we performed a robust clustering of various factors into different clusters of life satisfacton.



# Introduction

|   For this project, we utilized the OECD database focusing on **indicators and life-satisfaction scores across various regions within different countries**. This comprehensive dataset captures diverse dimensions of well-being, such as education, employment, health, environment, and social support, which together contribute to a region's overall quality of life. The indicators allow for meaningful comparisons between regions, highlighting disparities and trends. Although the data reflects different years depending on the country, for consistency we adopted the **latest available data**.

|   It is essential to acknowledge that while some variability exists in the timing of the data, we assumed that any changes within a year or two would be minimal and unlikely to significantly alter policies or life-satisfaction scores. This approach enables us to draw relevant conclusions about the factors influencing regional well-being. By examining key metrics such as income, safety, health, and life satisfaction, this report aims to present a clear picture of well-being across OECD regions and highlight patterns that can inform future policies and initiatives.

# Original Dataset

|   To execute this project, a dataset containing the different metrics regarding well-being was downloaded and linked to a variable named **df_wb**.

```{r, echo=FALSE}
df_wb <- read.xlsx("C:\\Users\\franc\\Documents\\GitHub\\ams_final_project\\Dataset\\OECD_Well_Being.xlsx")
```

|   The variable **df_wb** contains 447 tuples and 28 columns (25 are the attributes to analyze).
```{r, echo=FALSE}
dim(df_wb)
```

\newpage

|   The variables included are the following:

1.	**Country:** Includes the name of all the countries included in the results.
2.	**Region:** Includes the name of all the cities from the countries included in the results.
3.	**Code:** Code associated to each pair country-city.
4.	**Population.with.at.least.secondary.education.(%):** Percentage of the population completing secondary education.
5.	**Employment rate (%):** Percentage of the working-age population employed.
6.  **Unemploy-ment rate (%):** Percentage of people without jobs actively seeking work.
7.  **Household disposable income per capita (USD PPP):** Average income available per person, in USD.
8.  **Homicide rate (per 100k):** Number of homicides per 100.000 people.
9.  **Mortality rate (per 1k):** Deaths per 1,000 people annually.
10. **Life expectancy:**  Average expected lifespan (years).
11. **Air pollution (level of PM2.5, µg/m³):** Fine particulate air pollution levels.
12. **Voter turnout (%):** Share of voters participating in elections.
13. **Broadband access (% of household):** Percentage of households with internet access.
14. **Internet download speed 2021-Q4 (%):** Internet speed growth/decline in 2021-Q4.
15. **Number of rooms per person:** Average living space per person.
16. **Perceived social network support  (%):** Percentage of people with available social support.
17. **Self assessment of life satisfaction (0-10):** Subjective rating of overall happiness.
18. **Education (0-10):** Regional score for education.
19. **Jobs (0-10):** Score based on employment indicators.
20. **Income (0-10):** Score for household income.
21. **Safety (0-10):** Regional score for personal safety.
22. **Health (0-10):** Score for health indicators.
23. **Environment (0-10):** Score for environmental quality.
24. **Civic engagement (0-10):** Score for public participation.
25. **Accessibility to services (0-10):** Availability of public services.
26. **Housing (0-10):** Score for housing quality and affordability.
27. **Community (0-10):** Score for social cohesion.
28. **Life satisfaction (0-10):** Overall happiness score.

|   The value type of each attribute are the following:

```{r, echo=FALSE}
sapply(df_wb, class)
```

# Data Manipulation and EDA

## Null values

|   Before analyzing the hypothesis and attributes, a check on the data structure was conducted to prevent potential errors in the future. This involved examining both data types and null values.

|   First we checked the **null values** to determine their significance and understand which is the best action to take regaridng this matter. After checking that there were no null values, but yet we couldn't perform some calculations on the attributes, we did a more detailed analysis to realize that there were null values which were replaced by the **character ".."**.

```{r, echo=FALSE}
wb_missing_values <- sapply(df_wb, function(col) sum(col == "..")/nrow(df_wb)*100)
as.data.frame(wb_missing_values)
```

|   Taking this into account, we concluded that there were no significance level of null values (in the form of ".."), so for the moment we decided to keep all the information and look for further actions regarding null values.

```{r, echo=FALSE}
missing_values_country <- aggregate(. ~ Country+Region, data = df_wb, function(x) sum(x == ".."))
missing_values_country <- as.data.frame(missing_values_country)

missing_values_country <- aggregate(
  rowSums(df_wb[, 4:ncol(df_wb)] == "..") ~ Country + Region,
  data = df_wb,
  FUN = sum
)

colnames(missing_values_country) <- c("Country", "Region", "n_null")

missing_values_country <- missing_values_country[missing_values_country$n_null > 0, ]

missing_values_country <- table(missing_values_country$Country)
missing_values_country <- as.data.frame(missing_values_country)
colnames(missing_values_country) <- c("Country", "n_cities")

n_country <- table(df_wb$Country)
n_country <- as.data.frame(n_country)
colnames(n_country) <- c("Country", "n_cities")

missing_by_country <- merge(n_country, missing_values_country, by = "Country")
colnames(missing_by_country) <- c("Country", "n_cities", "n_missing")
missing_by_country$p_na <- (missing_by_country$n_missing/missing_by_country$n_cities)*100

missing_by_country
```

|   Also, we can see that, when analyzing the null values by country, there were some cases that have 100% of missing values in some attributes. This is why, we decided to drop **Costa Rica, Iceland and Japan** as they had attributes without values, and this would have not been useful for our project. In addition to them, we also decided to exclude Türkiye as we found poor the information contained in the Life.Satisfaction.(0-10) attribute.

|   Moreover, for those cases with some null values, we decided to replace them with the **median of the country**. We decided this because the data is divided by city, so we could use the median of the rest cities of the country for those cities with null values. Furthermore, the median is a better option than the mean as we can avoid the influence of any possible outlier.

|   Finally, after all this data manipulation, we can see that now the data type of each attribute is correct.

```{r, echo=FALSE}
df_wb <- df_wb[!(df_wb$Country == "Costa Rica" | df_wb$Country == "Iceland" | df_wb$Country == "Japan" | df_wb$Country == "Türkiye"), ]

df_wb <- df_wb %>% 
  mutate(across(everything(), ~ ifelse(. == "..", NA, .)))

df_wb <- df_wb %>%
  mutate(across(.cols = which(!names(df_wb) %in% names(df_wb)[1:3]), 
                .fns = ~ as.numeric(.)))

df_wb <- df_wb %>% 
  group_by(Country) %>% 
  mutate(across(3:(ncol(df_wb)-1), ~ replace_na(., median(., na.rm = TRUE)))) %>% 
  ungroup()

df_wb <- as.data.frame(df_wb)

sapply(df_wb, class)
```

## Variables and Correlation
|   Continuing with the data manipulation, we decided to carry out a correlation analysis to determine the need of any further removal of attributes. For this, we decided to plot a correlation heatmap to rapidly see any pair of attributes highly correlated, and with thise, determine the removal of one of them.

```{r, echo=FALSE}
cor_matrix <- cor(df_wb[,4:ncol(df_wb)], use = "complete.obs")

cor_data <- melt(cor_matrix)

ggplot(cor_data, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, 
                       limit = c(-1, 1), space = "Lab", 
                       name = "Correlation")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_blank(),
    axis.title.y = element_blank())+
  coord_fixed()
```

| As a result of this analysis, we decided to drop the following variables, because they can be explained by others (high correlation) and probably have less information than other correlated variables (this was checked manually with the dataset):

1.  **Unemploy-ment.rate.(%):** -0.9 of correlation with Jobs.(0-10).
2.  **Life.expectancy:** 0.9 of correlation with Health.(0-10).
3.  **Internet.download.speed.2021-Q4.(%):** 0.87 of correlation with Accessiblity.to.services.(0-10).
4.  **Perceived.social.network.support.(%):** 0.83 of correlation with Community.(0-10).
5.  **Voter.turnout.(%):** 0.99 of correlation with Civic.engagement.(0-10).
6.  **Air.pollution.(level.of.PM2.5,.µg/m³):** -0.97 of correlation with Environment.(0-10).
7.  **Population.with.at.least.secondary.education.(%):** 0.99 of correlation with Education.(0-10).
8.  **Household.disposable.income.per.capita.(USD.PPP):** 0.99 of correlation with Income.(0-10).
9.  **Employment.rate.(%):** 0.92 of correlation with Jobs.(0-10).

|   Additionally, we decided to remove also Homicide.rate.(per.100k) and Mortality.rate.(per.1k) because we assumed that can be represented by Safety (0-10). Also we removed Broadband.access.(%.of.household) and Number.of.rooms.per.person as we did not consider it useful for the project. Last but not least, we removed Self.assessment.of.life.satisfaction.(0-10) as we directly used Life.satisfaction.(0-10).
```{r, echo=FALSE}
df_wb <- df_wb[, c(1:3, 18:ncol(df_wb))]
```

|   Taking into account the resulting dataset, we decided to exclude those countries that have less than 10 cities in the dataset, as we have 10 variables plus the main variable (life satisfaction).

```{r, echo=FALSE}
df_wb <- df_wb %>%
  group_by(Country) %>%                        # Agrupar por país
  filter(n_distinct(Region) >= 10) %>%      # Contar ciudades únicas y filtrar
  ungroup() 
```

|   The last step of the data manipulation was to set the upper bound of the scale to 10, as there were some cases with decimals that ended up being a little bit over 10.

```{r, echo=FALSE}
df_wb[4:ncol(df_wb)] <- lapply(df_wb[4:ncol(df_wb)], function(x) ifelse(x > 10, 10, x))
```

|   So the final dataset used in the models was the following, containing 295 tuples and 14 columns (10 are the attributes to analyze, plus the life satisfaction attribute):

```{r, echo=FALSE}
dim(df_wb)

summary(df_wb)
```

\newpage

## EDA

|   After completing a rigorous data cleaning process to remove redundancies and address multicollinearity, we continued with an Exploratory Data Analysis (EDA) to obtain insights about the factors influencing life satisfaction across 15 countries.

### Life Satisfaction Distribution Across Countries

|   We began by analyzing how life satisfaction varies from one country to another. As illustrated in the box plot below, countries like the Netherlands, New Zealand, and Canada stood out with the highest average life satisfaction scores (9.36, 8.9, and 8.22, respectively). On the other hand, countries such as Greece, and Colombia reported the lowest scores.

|   Interestingly, in countries like Mexico and Colombia, we observed wide variability, suggesting significant regional differences within these nations. This led us to our **first key insight**: life satisfaction is not equally distributed across countries, and economic or social disparities likely play a role.

```{r, echo=FALSE}
#High-Level Distribution of Life Satisfaction

# Compare life satisfaction across countries

# Calculate both Average and Max Life Satisfaction for Annotation
country_avg <- df_wb %>%
  group_by(Country) %>%
  summarise(
    Average_Life_Satisfaction = mean(`Life.satisfaction.(0-10)`, na.rm = TRUE),
    Max_Life_Satisfaction = max(`Life.satisfaction.(0-10)`, na.rm = TRUE)
  )

# Boxplot with Average Annotations
ggplot(df_wb, aes(x = reorder(Country, `Life.satisfaction.(0-10)`), y = `Life.satisfaction.(0-10)`)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  geom_text(data = country_avg, 
            aes(x = Country, y = Max_Life_Satisfaction + 0.5, 
                label = round(Average_Life_Satisfaction, 2)), 
            size = 3, color = "red") +
  coord_flip() +
  ggtitle("Life Satisfaction by Country with Average Scores") +
  xlab("Country") +
  ylab("Life Satisfaction (0-10)") +
  theme_minimal()


```

### Understanding Relationships Between Metrics (Correlation Analysis)

|   Deeper in our analysis, we explored the correlations between life satisfaction and other well-being metrics using a heatmap:

1.    **Housing** had the strongest correlation with life satisfaction (0.70), showing that access to adequate housing and living conditions significantly influences well-being.

2.    **Jobs** metric followed closely at 0.6, highlighting the importance of employment opportunities.

3.    **Environemnt** and **Community** stood out with a correlation of 0.55 and 0.56 respectively, highlighting that a clean and sustainable environment and, connectedness and solidarity among groups in society are also an important part of life. 

|   Surprisingly, **Income** showed a slightly lower correlation at 0.54, suggesting that while important, economic wealth is not the sole driver of life satisfaction.

|   This analysis revealed that housing conditions, employment opportunities, environment and community are more influential for life satisfaction than income alone.

```{r, echo=FALSE}
# Heatmap of correlations

# Select only numeric columns for correlation
numeric_data <- df_wb %>% select_if(is.numeric)

# Compute the correlation matrix
cor_matrix <- cor(numeric_data, use = "complete.obs")



ggcorrplot(cor_matrix, 
           method = "square",       # Use square tiles for the heatmap
           type = "upper",          # Show only the upper triangle
           lab = TRUE,              # Display correlation values
           lab_size = 3,            # Size of the labels
           title = "Correlation Heatmap of Metrics", 
           colors = c("blue", "white", "red"))  # Custom color scale



```

\newpage

### A Holistic Comparison of Countries (Radar Chart)

|   To compare the overall performance of all 15 countries, we created a radar chart that visualizes average scores across various attributes such as Environment, Jobs, Housing, and Education. Countries like the Netherlands and New Zealand demonstrated balance across multiple dimensions, outstanding in areas such as Education, Jobs, and Environment.

|   Conversely, countries like Greece struggled across several metrics, particularly in Income, Health, and Safety, which correlates with their lower life satisfaction scores.
So from this visualization we can highlight some critical insight:
Top-performing countries achieve balance across economic, environmental, and social dimensions, while lower-performing countries face challenges in multiple areas.

```{r, echo=FALSE}
#Radar Chart for Top Countries

# Calculate average life satisfaction per country
country_ranking <- df_wb %>%
  group_by(Country) %>%
  summarise(Average_Life_Satisfaction = mean(`Life.satisfaction.(0-10)`, na.rm = TRUE)) %>%
  arrange(desc(Average_Life_Satisfaction))

# Selecting all countries (Top 15)
all_countries <- country_ranking$Country
all_data <- df_wb %>% filter(Country %in% all_countries)

# Aggregating mean attributes for each country
radar_data <- all_data %>%
  group_by(Country) %>%
  summarise_at(vars(`Education.(0-10)`, `Jobs.(0-10)`, `Income.(0-10)`, `Safety.(0-10)`,
                    `Health.(0-10)`, `Environment.(0-10)`, `Civic.engagement.(0-10)`,
                    `Accessiblity.to.services.(0-10)`, `Housing.(0-10)`, `Community.(0-10)`, 
                    `Life.satisfaction.(0-10)`), mean, na.rm = TRUE)

# Define Colors Dynamically for All Countries
country_colors <- distinctColorPalette(nrow(radar_data)) # Generate unique colors dynamically
names(country_colors) <- radar_data$Country # Assign country names to colors

# Clean column names to remove "(0-10)"
radar_data_clean <- radar_data %>%
  rename_with(~ str_replace(., "\\(0-10\\)", ""))

# getting the  Radar Chart Data ready 
max_vals <- rep(10, ncol(radar_data_clean) - 1)  # Set max values (10 for all attributes)
min_vals <- rep(0, ncol(radar_data_clean) - 1)   # Set min values (0 for all attributes)
radar_chart_data <- rbind(max_vals, min_vals, radar_data_clean[,-1])

# Adjust figure layout
par(mar = c(8, 4, 4, 2), xpd = TRUE, cex = 0.7) 

# Plot the Radar Chart WITHOUT Fill Colors
par(mar = c(8, 4, 4, 2), xpd = TRUE, cex = 0.8) 

radarchart(radar_chart_data, axistype = 1,
           pcol = country_colors[radar_data_clean$Country],  # Colors for outlines
           pfcol = NA,                                # Remove fill color
           plwd = 2, plty = 1,
           title = "Comparison of All 15 Countries Across Attributes")


# Adjusted Legend Below to make the Plot clear 
legend("bottom", inset = c(0, -0.4), 
       legend = radar_data_clean$Country,
       col = country_colors[radar_data_clean$Country],
       lty = 1, lwd = 3, cex = 0.7,
       ncol = 3, title = "Countries")

```

\newpage

### Visualizing Patterns Through Interactive Scatter Plots

|   Finally, we created two interactive plots to explore the relationships between Income, Civic Engagement, and Life Satisfaction:

1.    **Plot 1: Circle Size by Life Satisfaction, Color by Country:** This plot revealed that countries with larger circles (higher life satisfaction) tend to have higher income and civic engagement levels. For instance, the Netherlands and the United States dominate the upper-right region of the plot. Conversely, smaller circles in the lower-left region highlight countries like Greece, where income and civic engagement remain low.

```{r, echo=FALSE, fig.align='center', out.width='100%', warning=FALSE, message=FALSE}
# Interactive scatter plot: Income vs Life Satisfaction

# Custom color palette for countries
custom_colors <- c(
  "Canada" = "red", "Chile" = "orange", "Colombia" = "yellow", "France" = "blue", 
  "Germany" = "green", "Greece" = "purple", "Italy" = "cyan", "Lithuania" = "chocolate3", "Mexico" = "pink", 
  "Netherlands" = "navy", "New Zealand" = "darkgreen", "Poland" = "brown", 
  "Spain" = "violet", "United Kingdom" = "coral", "United States" = "darkred"
)


# Plot 1: Circle Size by Life Satisfaction, Color by Country

plot1 <- ggplot(df_wb, 
                aes(x = `Income.(0-10)`, 
                    y = `Civic.engagement.(0-10)`, 
                    size = `Life.satisfaction.(0-10)`, 
                    color = Country,  # Map color to Country
                    text = paste("Region:", Region, 
                                 "<br>Country:", Country, 
                                 "<br>Life Satisfaction:", `Life.satisfaction.(0-10)`))) +
  geom_point(alpha = 0.7) +  # Use only circles
  ggtitle("Plot 1: Circle Size by Life Satisfaction, Color by Country") +
  xlab("Income (0-10)") +
  ylab("Civic Engagement (0-10)") +
  theme_minimal() +
  scale_size_continuous(range = c(3, 12), name = "Life Satisfaction") +  # Adjust circle sizes
  scale_color_manual(values = custom_colors) +  # Custom country colors
  guides(color = guide_legend(title = "Country"), 
         size = guide_legend(title = "Life Satisfaction")) +
  theme(
    legend.position = "bottom",       # Move legend below the plot
    legend.box = "horizontal",        # Arrange legend items horizontally
    plot.margin = margin(10, 10, 30, 10), # Add extra margin at the bottom
    legend.text = element_text(size = 5), # Adjust legend text size
    legend.title = element_text(size = 5) # Adjust legend title size
  )

# Display Both Plots
plot1
```

\newpage

2.    **Plot 2: Different Shapes by Country, Color by Life Satisfaction:** Here, each country is represented with a unique shape, and colors reflect their life satisfaction. Countries with higher satisfaction are visibly clustered in red, whereas lower-satisfaction countries are more dispersed, particularly toward the lower income range. These visualizations allowed us to see the clear relationships between income, civic engagement, and well-being, while also identifying disparities and regional patterns.

```{r, echo=FALSE, fig.align='center', out.width='100%', warning=FALSE, message=FALSE}
# Interactive scatter plot: Income vs Life Satisfaction

# Custom color palette for countries
custom_colors <- c(
  "Canada" = "red", "Chile" = "orange", "Colombia" = "yellow", "France" = "blue", 
  "Germany" = "green", "Greece" = "purple", "Italy" = "cyan", "Lithuania" = "chocolate3", "Mexico" = "pink", 
  "Netherlands" = "navy", "New Zealand" = "darkgreen", "Poland" = "brown", 
  "Spain" = "violet", "United Kingdom" = "coral", "United States" = "darkred"
)


# Plot 2: Different Shapes by Country, Color by Life Satisfaction

plot2 <- ggplot(df_wb, 
                aes(x = `Income.(0-10)`, 
                    y = `Civic.engagement.(0-10)`, 
                    color = `Life.satisfaction.(0-10)`,  # Color by Life Satisfaction
                    shape = Country,                    # Shape by Country
                    text = paste("Region:", Region, 
                                 "<br>Country:", Country, 
                                 "<br>Life Satisfaction:", `Life.satisfaction.(0-10)`))) +
  geom_point(alpha = 0.7, size = 5) +  # Larger shapes for visibility
  ggtitle("Plot 2: Different Shapes by Country, Color by Life Satisfaction") +
  xlab("Income (0-10)") +
  ylab("Civic Engagement (0-10)") +
  theme_minimal() +
  scale_shape_manual(values = 0:20) +  # Unique shapes for up to 20 countries
  scale_color_gradient(low = "blue", high = "red") +  # Gradient color for Life Satisfaction
  guides(shape = guide_legend(title = "Country"), 
         color = guide_colorbar(title = "Life Satisfaction")) +
  theme(
    legend.position = "bottom",       # Move legend below the plot
    legend.box = "horizontal",        # Arrange legend items horizontally
    plot.margin = margin(10, 10, 30, 10), # Add extra margin at the bottom
    legend.text = element_text(size = 5), # Adjust legend text size
    legend.title = element_text(size = 5) # Adjust legend title size
  )

# Display Plot
plot2
```