---
title: "Regional Well Being"
author: Akash Mittal, Maged Saeed Abdo Mostafa Kharshom, Precious Prince, Franco Reinaldo
  Bonifacini
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(openxlsx)
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(fpp3)
library(GGally)
```

## Abstract

|   This project aims to **estimate the effect of various factors such as health, education, and income among others, on the life satisfaction of people in different regions from different countries**. For this, we carried out a robust approach with MEMs & Robust Estimators. Also, we performed a robust clustering of various factors into different clusters of life satisfacton.



## Introduction

|   For this project, we used an OECD database called **indicators and life-satisfaction scores for different regions within the different countries**. It is important to highlight that data is not constant regarding result's years among countries, but for the project's purpose it is constant as we considered it as **“the latest available data”**. We also assumed that in a  year or two, there are barely any drastic changes in a country policies or life-satisfaction scores.

```{r, echo=FALSE}
df_wb <- read.xlsx("C:\\Users\\franc\\Documents\\GitHub\\ams_final_project\\Dataset\\OECD_Well_Being.xlsx")
```

|   In the original dataset, which we called **df_wb**, we have 447 tuples and 28 columns.
```{r, echo=FALSE}
dim(df_wb)
```

|   The raw variables included are:

1.	**Country (chr):** includes the name of all the countries included in the results.
2.	**Region (chr):** includes the name of all the cities from the countries included in the results.
3.	**Code (chr):** code associated to each pair country-city.
4.	**Population.with.at.least.secondary.education.(%) (chr):** provides a certain freshness to the wine.
5.	