### Loading the Required Preliminary Libraries

```{r}
# Loading the required Libraries
library(dplyr)
library(ggplot2)
library(mclust)
library(readr)
library(openxlsx)
```

### Loading the Data from the Excel File and scaling it.
At the end we have 311 observations for 11 variables (+3 are Country, Region and Code)

```{r}
df <- read.xlsx("data_filtered.xlsx")
str(df)
df <- df %>%
  filter(Country != "Türkiye")
```

```{r}
# Selecting only the Numeric Columns (predictor columns) to scale the data

df_mod <- df %>% dplyr::select(-Country, -Region, -Code, -`Life.satisfaction.(0-10)`)

# Scaling the dataframe
df_mod <- scale(df_mod) # But gives a matrix as output

# Converting back to dataframe
df_mod <- as.data.frame(df_mod)

```


```{r}
# Checking the updated structure of the dataframe
str(df_mod)
```

### Performing the Model Based Clustering on the dataset.

### Clustering using MClust (Model Based Clustering)

```{r}
# Selecting 5 clusters and choosing an Optimal Cluster out of all the possible combinations 
wb_optimal_cst <- Mclust(df_mod, G=1:5) # setting the number of optimal clusters to be chosen from 1 to 5.

# Summary of Clusters with Model Fitted by EM algorithm and a view counts in each cluster along with BIC values. 
summary(wb_optimal_cst, parameters = TRUE)

```

The summary shows the number of optimal clusters to be 4. 
-> Mclust VEV (ellipsoidal, equal shape) model with 4 components. Fitted a Gaussian MIxture Model by EM Algorithm. 


###### Plot of BIC vs Number of Components Used for Clustering
###### Classification and Uncertainty plots of different clusters

```{r}

# Reference 1 - https://mclust-org.github.io/mclust/reference/plot.Mclust.html
# Reference 2 - https://mclust-org.github.io/mclust-book/chapters/03_cluster.html#:~:text=Model-based%20clustering%20%28MBC%29%20is%20a%20probabilistic%20approach%20to,described%20by%20a%20probability%20distribution%20with%20unknown%20parameters.

# Legend position (for styling)
legend_args <- list(x = "bottomright", ncol = 5)
```

```{r}
x11()
# Plotting BIC to assess model fit
plot(wb_optimal_cst, what = 'BIC', legendArgs = legend_args)
```

```{r}
x11()
# Plotting classification of data points
plot(wb_optimal_cst, what = 'classification', fillEllipses = TRUE)
```

```{r}
x11()
# Plotting the uncertainty 
plot(wb_optimal_cst, what = 'uncertainty')
```

```{r}
x11()
# Plot density distribution of the data within clusters
plot(wb_optimal_cst, what = 'density', legendArgs = legend_args)

```

```{r}
# Skipped Portion (Not necessary)
# uncertainty_mclust <- data.frame(
#   id = 1:nrow(df_mod),
#   cluster = wb_optimal_cst$classification,
#   uncertainty = wb_optimal_cst$uncertainty
# )
# 
# # Rescale uncertainty for better visualization (optional)
# uncertainty_mclust$rescaled_uncertainty <- uncertainty_mclust$uncertainty * 100  # Scaling by 100 for example
# 
# # Now, plot the rescaled uncertainty
# ggplot(uncertainty_mclust, aes(x = rescaled_uncertainty, y = reorder(id, rescaled_uncertainty))) +
#   geom_point() +
#   facet_wrap(~ cluster, scales = 'free_y', nrow = 1) +
#   theme_minimal()
```


###### Using PCA to plot the different clusters for visualization

```{r}
# Plotting the Four clusters using PC Components for better visualization.
# Performing PCA using prcomp
pca <- prcomp(df_mod, scale. = TRUE)

# first two principal components
pca_data <- data.frame(pca$x[, 1:2])

# Using the cluster assignments from Mclust
pca_data$Cluster <- factor(wb_optimal_cst$classification)

# Plotting the MClust clusters on PCA plot
ggplot(pca_data, aes(x = PC1, y = PC2, color = Cluster)) +
  geom_point(alpha = 0.9, size = 2) +
  theme_minimal() +
  labs(title = "PCA Plot with the 5 Clusters from MClust", x = "Principal Component 1", y = "Principal Component 2") +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d"))+
  theme(plot.title = element_text(hjust = 0.5))
```

###### Using T-SNE also for Visualization

```{r}
library(Rtsne)
set.seed(218)
# Performing t-SNE
tsne <- Rtsne(df_mod, dims = 2)

# t-SNE coordinates
tsne_data <- data.frame(tsne$Y)

# Cluster assignments from Mclust
tsne_data$Cluster <- factor(wb_optimal_cst$classification)

# Plotting t-SNE with MClust clusters
ggplot(tsne_data, aes(x = X1, y = X2, color = Cluster)) +
  geom_point(alpha = 0.9, size = 2) +
  theme_minimal() +
  labs(title = "t-SNE Plot with the 5 Clusters from MClust", x = "t-SNE Dimension 1", y = "t-SNE Dimension 2") +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}
# checking summary of df_mod just for an overview 
#summary(df_mod)
```

###### Adding the Clusters to original data frame and checking the means for each

```{r}
# Adding the cluster assignments to the original dataset
df_mod$Cluster <- factor(wb_optimal_cst$classification)

# Calculating  mean of each feature for every cluster and storing in dataframe cluster_means
cluster_means <- df_mod %>%
  group_by(Cluster) %>%
  summarise(across(everything(), mean))

# Viewing the cluster profiles
print(cluster_means)

```

###### Bar Plots for Each Feature and Cluster
This plot shows the contribution of each feature in the every cluster with base reference to the average (mean) value of the feature.
We see, for cluster 1, all the features have a value more than the average value for all parameters.

```{r}

library(tidyr)

# Convert the cluster_means to a long format for easy plotting
cluster_means_long <- cluster_means %>%
  pivot_longer(cols = -Cluster, names_to = "Feature", values_to = "Mean_Value")

# Creating the bar plot for each feature by cluster
ggplot(cluster_means_long, aes(x = Feature, y = Mean_Value, fill = Cluster)) +
  geom_bar(stat = "identity", position = "dodge", alpha=0.9,color = "black", linewidth = 0.1) +
  labs(title = "Average Feature Values by Cluster", 
       x = "Feature", 
       y = "Average Value") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  # Rotate x-axis labels
  scale_fill_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +  # Cluster colors
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

```


###### Heatmap displaying the comparison for each cluster and the values - Skipped

```{r}
# # Cross-tabulating the original life satisfaction clusters with predicted clusters
# cluster_comparison <- table(df$`Life.satisfaction.(0-10)`, df_mod$Cluster)
# 
# # Create a heatmap of the comparison
# library(pheatmap)
# pheatmap(cluster_comparison, 
#          color = colorRampPalette(c("white", "blue"))(100),
#          main = "Comparison Heatmap",
#          display_numbers = TRUE,
#          cluster_rows = TRUE, 
#          cluster_cols = TRUE,
#          angle_col = 45)

```

###### Uncertainty Plot of the 4 MClust clusters

```{r}
# Plot uncertainty of the cluster assignments
plot(wb_optimal_cst, what = "uncertainty")

```

###### Scatter plot of Income vs Civic Engagement for the 4 Clusters
To show the distribution of the four clusters only.

```{r}
# Create a scatter plot for Income vs Civic Engagement with clusters
ggplot(df_mod, aes(x = `Income.(0-10)`, y = `Civic.engagement.(0-10)`, color = as.factor(wb_optimal_cst$classification))) +
  geom_point(alpha = 0.9, size = 2) +
  labs(title = "Income vs Civic Engagement (Scaled) by Cluster", 
       x = "Income", y = "Civic Engagement", 
       color = "Cluster") +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +  # Adjust color palette as needed
  theme_minimal()+
  theme(plot.title = element_text(hjust=0.5))

```
###### Plot with the centroids for the Income vs Civic Engagement for the 4 Clusters

```{r}

# Calculating the centroids of clusters (average value of Income and Civic Engagement for each cluster)

centroids <- df_mod %>%
  group_by(Cluster) %>%
  summarise(Income_mean = mean(`Income.(0-10)`),
            Civic_mean = mean(`Civic.engagement.(0-10)`))

# Plotting the scatter plot with centroids
ggplot(df_mod, aes(x = `Income.(0-10)`, y = `Civic.engagement.(0-10)`, color = as.factor(wb_optimal_cst$classification))) +
  geom_point(alpha = 0.9, size = 2) +
  geom_point(data = centroids, aes(x = Income_mean, y = Civic_mean), 
             color = "grey1", size = 4, shape = 8, stroke = 2) +  # Centroids as large black stars
  labs(title = "Income vs Civic Engagement (Scaled) by Cluster with Centroids", 
       x = "Income", y = "Civic Engagement", 
       color = "Cluster") +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

```
###### Cluster Profile Heatmap with average values of factors in each cluster

The values are from base line 0, as the data had been scaled.
-> This plot is shown in presentation in combination with the Average Plot.

```{r}

library(pheatmap)

# Calculate the mean of each feature by cluster
cluster_means <- df_mod %>%
  group_by(Cluster) %>%
  summarise(across(starts_with("Education"):starts_with("Community"), mean))

# Plot heatmap
pheatmap(as.matrix(cluster_means[, -1]), 
         cluster_rows = TRUE, cluster_cols = TRUE,
         main = "Cluster Profiles Heatmap",
         scale = "row",  # Scale variables by row (features)
         display_numbers = TRUE)

```

###### Bar Plot for all the 4 clusters showing the average Life Satisfaction Score in Each cluster

It shows the average life satisfaction score in each cluster.

```{r}

# Adding the predicted clusters to the original dataset
df$Cluster <- wb_optimal_cst$classification

# Calculating the mean Life Satisfaction score for each cluster
life_satisfaction_avg <- df %>%
  group_by(Cluster) %>%
  summarise(Average_Life_Satisfaction = mean(`Life.satisfaction.(0-10)`))

# View the results
print(life_satisfaction_avg)

# Visualizing using a bar plot of average Life Satisfaction by cluster
ggplot(life_satisfaction_avg, aes(x = as.factor(Cluster), y = Average_Life_Satisfaction, fill = as.factor(Cluster))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(Average_Life_Satisfaction, 2)),  # Adding labels (rounded to 2 decimal places)
            vjust = -0.5,  # Adjust position of text above the bar
            color = "black") +  # Color of the text
  labs(title = "Average Life Satisfaction by Cluster", x = "Cluster", y = "Average Life Satisfaction (0-10)") +
  scale_fill_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +  # Adjust colors for clusters
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(hjust=0.5))
```

```{r}
# str(df)
```

###### Plotting Histograms and Density plots for the 4 clusters

```{r}
# Density Plot for Life.Satisfaction Score
# histogram(df$`Life.satisfaction.(0-10)`, df, type = c("count"))
# densityplot(df$`Life.satisfaction.(0-10)`, df)
```

Histogram

```{r}

# Creating histogram and highlighting the points by 4 cluster
ggplot(df, aes(x = `Life.satisfaction.(0-10)`, fill = as.factor(Cluster))) +
  geom_histogram(bins = 30, color = "black", alpha = 0.7) +
  labs(title = "Histogram of Life Satisfaction by Cluster", 
       x = "Life Satisfaction (0-10)", 
       y = "Frequency") +
  scale_fill_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +  # Adjust colors for clusters
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))


```

Density Plots for the 4 clusters.

```{r}

# Create a density plot and highlight the clusters
ggplot(df, aes(x = `Life.satisfaction.(0-10)`, color = as.factor(Cluster))) +
  geom_density(size = 1.5) +  # Adjust the line thickness
  labs(title = "Density Plot of Life Satisfaction by Cluster", 
       x = "Life Satisfaction (0-10)", 
       y = "Density") +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +  # Adjust colors for clusters
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

```

```{r}

```


##################

## Using K-Means for Clustering and comparing the Clusters

### Doing K-Means Clustering to cross check

```{r}
# Creating a new dataframe for K-Means
df_mod_k <- df_mod %>% select(-Cluster)
# str(df_mod_k)
```

###### Running K-Means

```{r}

# K-means clustering for 10 different values of k=1 to k=10.
set.seed(123)  # Seed for reproducibility
kmeans_results <- list()
for (k in 1:10) {
  kmeans_results[[k]] <- kmeans(df_mod_k, centers = k, nstart = 100)
}


```

###### Elbow Method For Number of Cluster Selection
It shows a kink at K=2, but we move ahead to choose upto 4 clusters after which the Within-Cluster sum of squares starts to fade off.

```{r}

# Plotting the total within-cluster sum of squares for different k values
wss <- sapply(kmeans_results, function(x) x$tot.withinss)
plot(1:10, wss, type = "b", pch = 19, xlab = "Number of Clusters", 
     ylab = "Total Within-Cluster Sum of Squares", main = "Elbow Method for K-means")

```

###### Running K-Means with 4 Clusters

```{r}
# Running K-means, k=4
kmeans_final <- kmeans(df_mod_k, centers = 4, nstart = 100)

# Adding K-means cluster assignment to the dataframe
df_mod_k$Cluster_Kmeans <- as.factor(kmeans_final$cluster)

# Overview of the cluster assignments in 5 clusters
cat("Count of Cities in Each Cluster from K-Means\n")
table(df_mod_k$Cluster_Kmeans)


```

### Comparing K-Means and MCluSt

```{r}
# Clusters from both the Models
df_mod_k$Cluster_Mclust <- factor(wb_optimal_cst$classification) # Mclust
df_mod_k$Cluster_Kmeans <- as.factor(kmeans_final$cluster) # K-Means

```

###### Contingency Table for the Two Models

```{r}
# Contingency table comparing Mclust and K-means cluster assignments
cluster_comparison <- table(df_mod_k$Cluster_Mclust, df_mod_k$Cluster_Kmeans)

cluster_comparison_r_c_sum <- addmargins(cluster_comparison)

# Printing the comparison
cat("\nContingency Table: Mclust and K-means Clusters Comparison\n")
cat("\nRows: Mclust Clusters\nColumns: K-means Clusters\n")
print(cluster_comparison_r_c_sum)
```

##### Scatter Plot Comparing the Clusters by K-Means and MClust

```{r}
# Scatter plot comparing K-means and Mclust clustering results
ggplot(df_mod_k, aes(x = Cluster_Mclust, y = Cluster_Kmeans, color = Cluster_Mclust)) +
  geom_jitter(alpha = 0.9, size=2) +
  labs(title = "Comparison of Mclust and K-means Clustering",
       x = "Mclust Clusters", y = "K-means Clusters") +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))
```
## Further, Analysis

###### Country Wise Distribution across different clusters
```{r}

# Adding the Country Column to df_mod
df_mod$Country <- df$Country

# Summary table with count of the number of entries for each country in each cluster
country_cluster_distribution <- df_mod %>%
  group_by(Cluster, Country) %>%
  summarise(Count = n(), .groups = 'drop')
```

###### Plot of Country Wise Distribution across the Clusters
This plot shows how the different cities of the countries are segregated/ distributed in the 4 clusters

```{r}

# Cluster-Wise Distribution of Countries

library(RColorBrewer)

custom_palette_15 <- c("#004949","#009292", "#490092","#006ddb","#b66dff","#F6D854", "#E41A1C", "#4DAF4A", "#A65628", "#f1cbcc",
                       "#999999", "#FC8D62", "#8DA0CB", "#E78AC3", "#A6D854")

ggplot(country_cluster_distribution, aes(x = Cluster, y = Count, fill = Country)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Count), position = position_stack(vjust = 0.5), color = "white") +  # Counts Inside the Horizontal Strips
  coord_flip() +  # Flipping Axes For horizontal bars
  labs(title = "Cluster-wise Distribution of Countries", 
       x = "Cluster", y = "Count of Cities from Each Country") +
  scale_fill_manual(values = custom_palette_15) +  # Using the custom color palette
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```
###### Rand Index Comparison

```{r}

adjustedRandIndex(wb_optimal_cst$classification, kmeans_final$cluster)

```

Thus, shows a strong agreement between the two clustering solutions i.e. a good overlap of the clusters by both techniques.

##################################################################

### Further check by specifically dividing the groups into 4 categories and comparing the clusters.

### Need to Do analysis of Original Clusters vs CLusters as Identified by Mclust and K-Means

## Analysis By dividing the Life.Satisfaction Score into 4 groups 0-10 divided into 2 groups of 2 scale each.

###### Dividing the Life.satisfaction.(0-10) score into 4 clusters-

```{r}

# Creating 4 clusters based on Life Satisfaction scores
df$LifeSatisfactionCluster <- cut(df$`Life.satisfaction.(0-10)`, 
                                  breaks = c(0, 2.5, 5, 7.5, 10), 
                                  labels = c("0-2.5", "2.5-5", "5-7.5", "7.5-10"), 
                                  include.lowest = TRUE)

# Checking the new column
table(df$LifeSatisfactionCluster)

```

###### Comparison with M-Clust

Contingency Table

```{r}

# Creating contingency table
mclust_clusters <- wb_optimal_cst$classification  # wb_optimal_cst contains Mclust clusters
contingency_table_mclust <- table(df$LifeSatisfactionCluster, mclust_clusters)

# Viewing the table
print(contingency_table_mclust)

```

Heatmap

```{r}

library(ggplot2)
library(reshape2)

# Melt the contingency table for plotting
melted_table <- melt(contingency_table_mclust)

# Rename the columns for clarity
colnames(melted_table) <- c("LifeSatisfactionCluster", "MclustCluster", "value")

# Heat Map
ggplot(data = melted_table, aes(x = MclustCluster, y = LifeSatisfactionCluster, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "lightblue3") +
  labs(title = "Heatmap of Life Satisfaction vs Mclust Clusters", 
       x = "Mclust Clusters", 
       y = "Life Satisfaction Clusters") + 
  theme_minimal()+
  theme(plot.title = element_text(hjust=0.5))
```

Visualization with PCA

```{r}
# Perform PCA on df_mod (standardized data)
pca_res <- prcomp(df_mod[, -c(11, 12)], scale. = TRUE)

# Add cluster labels to PCA results
pca_data <- data.frame(pca_res$x, 
                       LifeSatisfactionCluster = as.factor(df$LifeSatisfactionCluster), 
                       MclustCluster = as.factor(mclust_clusters))

# PCA Plot with Manual Colors for LifeSatisfactionCluster
ggplot(pca_data, aes(x = PC1, y = PC2, color = LifeSatisfactionCluster, shape = MclustCluster)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) + # Colors for LifeSatisfactionCluster
  scale_shape_manual(values = c(16, 17, 18, 15, 8)) +  # Different shapes for Mclust Clusters
  labs(title = "PCA Comparison of Life Satisfaction and Mclust Clusters",
       x = "Principal Component 1", 
       y = "Principal Component 2",
       color = "Life Satisfaction Clusters", 
       shape = "Mclust Clusters") +
  theme_minimal()

```

The plot shows the comparison of Mclust Clusters with the Original Clusters.

#### K-Means Cluster Comparison with Life-Satisfaction Original Clusters

Contingency Table

```{r}

# Assuming kmeans_final$cluster contains K-Means cluster assignments
contingency_table_kmeans <- table(df$LifeSatisfactionCluster, kmeans_final$cluster)

# Viewing the table
print(contingency_table_kmeans)

```

Heatmap

```{r}

# Melt the contingency table for plotting
melted_table_kmeans <- melt(contingency_table_kmeans)

# Heat Map for K-Means Clusters
ggplot(data = melted_table_kmeans, aes(x = as.factor(Var1), y = as.factor(Var2), fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "pink3") +
  labs(title = "Heatmap of Life Satisfaction vs K-Means Clusters", 
       x = "K-Means Clusters", 
       y = "Life Satisfaction Clusters",
       fill = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

```

Visualization

```{r}
# Add K-Means cluster labels to PCA results
pca_data$KMeansCluster <- as.factor(kmeans_final$cluster)

# PCA Plot with Manual Colors and Shapes for K-Means
ggplot(pca_data, aes(x = PC1, y = PC2, color = LifeSatisfactionCluster, shape = KMeansCluster)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) + # Colors for Life Satisfaction Clusters
  scale_shape_manual(values = c(16, 17, 18, 15, 8)) +  # Different shapes for K-Means Clusters
  labs(title = "PCA Comparison of Life Satisfaction and K-Means Clusters",
       x = "Principal Component 1", 
       y = "Principal Component 2",
       color = "Life Satisfaction Clusters", 
       shape = "K-Means Clusters") +
  theme_minimal()
```

The plot shows the comparison of K-Means Clusters with the Original Clusters.

Note: The clusters when divided manually into 4 groups, they don't overlap the clusters made by K-Means and Mclust. The probable reason being that the scores are subjective and different people measure life satisfaction scores in different perspectives, for some income would be a major factor, and for others community would be a major factor.
However, statistically, we see that the clustering using Mclust and K-Means show a good overlap for the 4 clusters with a rand index 0.80.

##############################################################################

### Robust K-Means / K-Medoids

```{r}
library(cluster)

# K-medoids 
kmedoids_res <- pam(df_mod, k = 4)  # 4 clusters

# Clustering results
table(kmedoids_res$clustering)

# Plot of the clusters
ggplot(df_mod, aes(x = `Income.(0-10)`, y = `Civic.engagement.(0-10)`, color = as.factor(kmedoids_res$clustering))) +
  geom_point(alpha = 0.8) +
  labs(title = "K-medoids (Robust K-Means)", x = "Income", y = "Civic Engagement", color = "Cluster") +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898")) +  
  theme_minimal() +
  theme(plot.title = element_text(hjust="0.5"))

```

##### Contingency Table for K-Medoids and Mclust

```{r}

kmedoids_clusters <- kmedoids_res$clustering


# Contingency table comparing K-Medoids and Mclust cluster assignments
cluster_comparison_kmedoids_mclust <- table(df_mod_k$Cluster_Mclust, kmedoids_clusters)

# Adding row and column sums to the contingency table
cluster_comparison_r_c_sum_kmedoids_mclust <- addmargins(cluster_comparison_kmedoids_mclust)

# Printing the comparison
cat("\nContingency Table: K-Medoids and Mclust Clusters Comparison\n")
cat("\nRows: Mclust Clusters\nColumns: K-Medoids Clusters\n")
print(cluster_comparison_r_c_sum_kmedoids_mclust)

```

###### Scatter Plot

```{r}

# Scatter plot comparing K-medoids and Mclust clustering results
ggplot(df_mod_k, aes(x = Cluster_Mclust, y = kmedoids_clusters, color = as.factor(kmedoids_clusters))) +
  geom_jitter(alpha = 0.9, size=2) +
  labs(title = "Comparison of Mclust and K-medoids Clustering",
       x = "Mclust Clusters", y = "K-medoids Clusters") +
  scale_color_manual(values = c("#d0eab8", "#eec77d", "#ce6d1d", "#ce9898", "#740d0d")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

```


### Comparison Metrics

##### Adjusted Rand INdex


```{r}

# Compute Adjusted Rand Index (ARI)
ari_kmeans_pam <- adjustedRandIndex(kmeans_final$cluster, kmedoids_res$clustering)
ari_kmeans_mclust <- adjustedRandIndex(kmeans_final$cluster, wb_optimal_cst$classification)
ari_pam_mclust <- adjustedRandIndex(kmedoids_res$clustering, wb_optimal_cst$classification)

```

###### Output of Rand Indexes

```{r}
# Print the results
print(paste("ARI between K-Means and K-Medoids:", ari_kmeans_pam))
print(paste("ARI between K-Means and Mclust:", ari_kmeans_mclust))
print(paste("ARI between K-Medoids and Mclust:", ari_pam_mclust))
```


We see that there is a only a minor agreement between clusters made from K-Means and K-Medoids, which is due to the reason that K-Medoids is more robust, uses representative points and doesn't work on centroids of data.
Further, we see a strong agreement between MClust and K-Means, which means data is probably in a structure suitable for both models. 
Also, K-Medoids is a distance based appraoch, while MClust is probabilistic approach.
