---
title: "AMS_Well_Being"
author: "Akash Mittal"
date: "2024-12-15"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of Estimation of Different factors on Life Satisfaction Score for different countries, using OECD survey dataset.

###### Loading Necessary Libraries

```{r}
# Loading the required libraries

library(openxlsx) # to read and write Excel files (.xlsx) in R
library(tidyr) # data format conversion, tidying and transforming data
library(tidyverse) # Whole package of libraries (dplyr, tidyr, ggplot2, readr, purrr, tibble, stringr, forcats)

library(reshape2) # for reshaping and transforming data
library(fpp3) # used for forecasting in R
library(GGally) # functionality of ggplot2 as it provides additional geoms and functions for data visualization

library(lme4) # for linear mixed effects modeling + functions for fitting and analyzing linear mixed effects models
library(insight) # functions for model-agnostic interpretation and visualization of regression models

```

###### Loading the dataset

```{r}

data = read.xlsx("OECD_Well_Being.xlsx")

```

###### Reading ans displaying the file

```{r}

str(data) # We have 447 tuples and 14 columns (25 are the attributes to analyze)

```
###### Concerting the special character .. in data for NUlls to NA

```{r}

data[data == ".."] <- NA

```


###### Converting the datatypes

```{r}

data <- data %>%
  dplyr::mutate(
    `Education.(0-10)` = base::as.numeric(as.character(`Education.(0-10)`)),
    `Jobs.(0-10)`= base::as.numeric(as.character(`Jobs.(0-10)`)),
    `Income.(0-10)`= base::as.numeric(as.character(`Income.(0-10)`)),
    `Safety.(0-10)`= base::as.numeric(as.character(`Safety.(0-10)`)),
    `Health.(0-10)`= base::as.numeric(as.character(`Health.(0-10)`)),
    `Environment.(0-10)`= base::as.numeric(as.character(`Environment.(0-10)`)),
    `Civic.engagement.(0-10)`= base::as.numeric(as.character(`Civic.engagement.(0-10)`)),
    `Accessiblity.to.services.(0-10)`=base::as.numeric(as.character(`Accessiblity.to.services.(0-10)`)),
    `Housing.(0-10)`= base::as.numeric(as.character(`Housing.(0-10)`)),
    `Community.(0-10)`= base::as.numeric(as.character(`Community.(0-10)`)),
    `Life.satisfaction.(0-10)`= base::as.numeric(as.character(`Life.satisfaction.(0-10)`))
  )

```


###### Checking Nulls

```{r}
data_na <- colSums(is.na(data))

  # sum(
  #   length(
  #     which(
  #       is.na(x)
  #       )
  #     )
  #   )

```

###### Data Summary

```{r}
summary(data)
```

###### Grouping data by country

```{r}

missing_values_country <- data %>%
  group_by(Country) %>%
  summarise(
    Education = sum(is.na(`Education.(0-10)`)),
    Jobs = sum(is.na(`Jobs.(0-10)`)),
    Income = sum(is.na(`Income.(0-10)`)),
    Safety = sum(is.na(`Safety.(0-10)`)),
    Health = sum(is.na(`Health.(0-10)`)),
    Environment = sum(is.na(`Environment.(0-10)`)),
    `Civic.engagement.(0-10)` = sum(is.na(`Civic.engagement.(0-10)`)),
    `Accessiblity.to.services.(0-10)` = sum(is.na(`Accessiblity.to.services.(0-10)`)),
    Housing = sum(is.na(`Housing.(0-10)`)),
    Community = sum(is.na(`Community.(0-10)`)),
    `Life.satisfaction.(0-10)` = sum(is.na(`Life.satisfaction.(0-10)`))
  )

```

```{r}
view(missing_values_country)
```


### Let's impute the missing value by a median value for that country

```{r}

country_data_count <- data %>%
  group_by(Country) %>%
  count()

view(country_data_count)

```

Since all the 10 values in Japan, and all 2 values in Iceland are missing, we are skipping Japan & Iceland from our analysis

```{r}

data_md <- data %>%
  dplyr::filter(Country != 'Japan' & Country!= 'Iceland')

```

Let's impute the remaining values using the Medians for that country

```{r}

data_md <- data_md %>%
  group_by(Country) %>%
  dplyr::mutate(across(
    c(
      `Education.(0-10)`,
      `Jobs.(0-10)`,
      `Income.(0-10)`,
      `Safety.(0-10)`,
      `Health.(0-10)`,
      `Environment.(0-10)`,
      `Civic.engagement.(0-10)`,
      `Accessiblity.to.services.(0-10)`,
      `Housing.(0-10)`,
      `Community.(0-10)`,
      `Life.satisfaction.(0-10)`
      ),
    ~ ifelse(is.na(.), median(., na.rm=TRUE), .)
  )
  )

```

Also, lets make the upper limit to 10, and replace any value >10 to 10

```{r}

data_md <- data_md %>%
  mutate(across(
    where(is.numeric),
    ~ pmin(.,10) # checks pariwise minimum value comparing with 10, and if value>10, replaces by 10
  ))
  

```


###### Dataframe with no Missing values (data_md), all replaced by median.

```{r}

# Filtering the countries with more than 10 cities 

data_filtered <- data_md %>%
  group_by(Country) %>%
  filter(n()>10) %>%
  ungroup()

```

Data Structure ->

```{r}

str(data_filtered)

```

```{r}

summary(data_filtered)

```

```{r}
# Also filtering out Turkey as data for Turkey is inconsistent
data_filtered <- data_filtered %>%
  filter(Country != "Türkiye")
```


```{r}
str(data_filtered)
```

##### Now, let's plot our data and see how the values are distributed.

```{r}
#x11()

p <- ggplot(data_filtered, aes(x = Country, y = `Health.(0-10)`)) +
  geom_boxplot(aes(fill = Country), outlier.shape = 18, outlier.size = 1) +
  #labs(title = "Health Scores Across Countries", x = "Country", y = "Health (0-10)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal()

a <- plotly::ggplotly(p)
a
```



```{r}
# Plotting the scatter plots of all numeric columns 

data_numeric <- data_filtered %>% select(where(is.numeric))

# Create the pair plot
ggpairs(data_numeric) 

```




```{r}
# Let's take log transformations of all the values in data set, to make and see how the data plots look
# Not used in analysis was just used to test.

data_filtered_lg <- data_filtered %>%
  mutate(across(
    where(is.numeric), ~ log(.)))

# Create the pair plot
#ggpairs(data_filtered_lg %>% select(where(is.numeric))) 

```


```{r}
summary(data_filtered)
```

###############################################################################
Exploratory Data Analysis

```{r}

#High-Level Distribution of Life Satisfaction

# Compare life satisfaction across countries

# Calculate both Average and Max Life Satisfaction for Annotation
country_avg <- data_filtered %>%
  group_by(Country) %>%
  summarise(
    Average_Life_Satisfaction = mean(`Life.satisfaction.(0-10)`, na.rm = TRUE),
    Max_Life_Satisfaction = max(`Life.satisfaction.(0-10)`, na.rm = TRUE)
  )

# Boxplot with Average Annotations
ggplot(data_filtered, aes(x = reorder(Country, `Life.satisfaction.(0-10)`), y = `Life.satisfaction.(0-10)`)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  geom_text(data = country_avg, 
            aes(x = Country, y = Max_Life_Satisfaction + 0.5, 
                label = round(Average_Life_Satisfaction, 2)), 
            size = 3, color = "red") +
  coord_flip() +
  ggtitle("Life Satisfaction by Country with Average Scores") +
  xlab("Country") +
  ylab("Life Satisfaction (0-10)") +
  theme_minimal()

```



### Heatmap of Correlations

```{r}
library(corrplot)

# Select only numeric columns for correlation
numeric_data <- data_filtered %>% select_if(is.numeric)

# Compute the correlation matrix
cor_matrix <- cor(numeric_data, use = "complete.obs")

# Customize the heatmap with better formatting
corrplot(cor_matrix, 
         method = "color",       # Use color gradient
         type = "upper",         # Show upper triangle
         addCoef.col = "black",  # Add correlation coefficients in black
         tl.col = "black",       # Text labels color
         tl.cex = 0.59,           # Text labels size
         number.cex = 0.59)       # Coefficient size)

```


### A Holistic Comparison of Countries (Radar Chart)

The radar chart highlights the Top 3 countries (Netherlands, Canada, and New Zealand) and the Bottom 3 countries (Poland, Greece, and Colombia) based on key life attributes. Netherlands leads in Jobs, Civic Engagement, and Accessibility to Services, achieving the highest Life Satisfaction. Canada excels in Housing, Environment, and Income, demonstrating a strong economic and living standard. New Zealand stands out with the best Community scores, fostering a strong sense of belonging.

Among the lowest countries, Poland performs well in Education but struggles with Environment and Housing. Colombia, despite having a better Environment than expected, lags in Housing, Accessibility to Services, and especially Safety, making it the least secure. This highlights how gaps in infrastructure, safety, and education can lead to lower life satisfaction across these countries.

```{r}

# Calculate average life satisfaction and filter Top 3 & Bottom 3 countries
country_ranking <- data_filtered %>%
  group_by(Country) %>%
  summarise(Average_Life_Satisfaction = mean(`Life.satisfaction.(0-10)`, na.rm = TRUE)) %>%
  arrange(desc(Average_Life_Satisfaction))

selected_countries <- c(
  head(country_ranking$Country, 3),  # Top 3 countries
  tail(country_ranking$Country, 3)   # Bottom 3 countries
)

# Aggregate radar chart data for the selected countries
radar_data <- data_filtered %>%
  filter(Country %in% selected_countries) %>%
  group_by(Country) %>%
  summarise_at(vars(`Education.(0-10)`, `Jobs.(0-10)`, `Income.(0-10)`, `Safety.(0-10)`,
                    `Health.(0-10)`, `Environment.(0-10)`, `Civic.engagement.(0-10)`,
                    `Accessiblity.to.services.(0-10)`, `Housing.(0-10)`, `Community.(0-10)`, 
                    `Life.satisfaction.(0-10)`), mean, na.rm = TRUE)

# Prepare radar chart data
max_vals <- rep(10, ncol(radar_data) - 1)  # Max scaling
min_vals <- rep(0, ncol(radar_data) - 1)   # Min scaling
radar_chart_data <- rbind(max_vals, min_vals, radar_data[,-1])

# Define vibrant colors for Top and Bottom countries
country_colors <- c("red", "blue", "green", "purple", "orange", "darkred")  # Custom strong colors
names(country_colors) <- radar_data$Country

library(fmsb) # used for creating radar charts

par(mar = c(2, 2, 2, 2), xpd = TRUE)

radarchart(radar_chart_data, axistype = 1,
           pcol = country_colors[radar_data$Country],
           pfcol = scales::alpha(country_colors[radar_data$Country], 0.2),
           plwd = 3, plty = 1,
           cglcol = "grey", cglty = 1, cglwd = 0.8, vlcex = 0.7,
           title = "Top 3 and Bottom 3 Countries - Radar Chart")

legend("bottomleft", legend = radar_data$Country,
       col = country_colors[radar_data$Country], lty = 1, lwd = 3, cex = 0.8, 
       title = "Countries", bg = "white", box.lwd = 0)


```

### Visualizing Patterns Through Scatter Plots

Finally, we created two plots to explore the relationships between Income, Civic Engagement, and Life Satisfaction:

1. Plot 1: Circle Size by Life Satisfaction, Color by Country: This plot revealed that countries with larger circles (higher life satisfaction) tend to have higher income and civic engagement levels. For instance, the Netherlands and the United States dominate the upper-right region of the plot. Conversely, smaller circles in the lower-left region highlight countries like Greece, where income and civic engagement remain low.

```{r}
# Custom color palette for countries
custom_colors <- c(
  "Canada" = "red", "Chile" = "orange", "Colombia" = "yellow", "France" = "blue", 
  "Germany" = "green", "Greece" = "purple", "Italy" = "cyan", "Lithuania" = "chocolate3", "Mexico" = "pink", 
  "Netherlands" = "navy", "New Zealand" = "darkgreen", "Poland" = "brown", 
  "Spain" = "violet", "United Kingdom" = "coral", "United States" = "darkred"
)


# Plot 1: Circle Size by Life Satisfaction, Color by Country

plot1 <- ggplot(data_filtered, 
                aes(x = `Income.(0-10)`, 
                    y = `Civic.engagement.(0-10)`, 
                    size = `Life.satisfaction.(0-10)`, 
                    color = Country,  # Map color to Country
                    text = paste("Region:", Region, 
                                 "<br>Country:", Country, 
                                 "<br>Life Satisfaction:", `Life.satisfaction.(0-10)`))) +
  geom_point(alpha = 0.7) +  # Use only circles
  ggtitle("Plot 1: Circle Size by Life Satisfaction, Color by Country") +
  xlab("Income (0-10)") +
  ylab("Civic Engagement (0-10)") +
  theme_minimal() +
  scale_size_continuous(range = c(3, 12), name = "Life Satisfaction") +  # Adjust circle sizes
  scale_color_manual(values = custom_colors) +  # Custom country colors
  guides(color = guide_legend(title = "Country"), 
         size = guide_legend(title = "Life Satisfaction")) +
  theme(
    legend.position = "bottom",       # Move legend below the plot
    legend.box = "horizontal",        # Arrange legend items horizontally
    plot.margin = margin(10, 10, 30, 10), # Add extra margin at the bottom
    legend.text = element_text(size = 5), # Adjust legend text size
    legend.title = element_text(size = 5) # Adjust legend title size
  )

# Display Both Plots
plot1
```



###############################################################################
# Fitting the Regression Models

#Model 1:Linear Regression

```{r}

fit_model_1 = lm(`Life.satisfaction.(0-10)` ~ ., data = data_filtered %>% select(-Country, -Region, -Code))
summary(fit_model_1)
```
The basic regression model shows that Jobs, Community, Environment, and Housing are significant for the Life satisfaction score. 


#Model 2:Linear Regression using Pairwise Interaction effects

```{r}

fit_model_2 <- lm(`Life.satisfaction.(0-10)` ~ .^2, data = data_filtered %>% select(-Country, -Region, -Code))
summary(fit_model_2)

```

But we observe that some variables have might have interaction effects, so we build a different model, and compare with the original one using ANOVA.

# Model 3: Model with Interaction effects based on relevance of factors.

```{r}

fit_model_3 <- lm(`Life.satisfaction.(0-10)` ~ `Education.(0-10)` + `Jobs.(0-10)` +
                  `Income.(0-10)` + `Safety.(0-10)` + `Health.(0-10)` +
                  `Environment.(0-10)` + `Civic.engagement.(0-10)` +
                  `Accessiblity.to.services.(0-10)` + `Housing.(0-10)` + 
                  `Community.(0-10)` + 
                  `Education.(0-10)`:`Safety.(0-10)` + `Education.(0-10)`:`Accessiblity.to.services.(0-10)` + `Education.(0-10)`:`Housing.(0-10)`+
                  `Jobs.(0-10)`:`Income.(0-10)` + `Jobs.(0-10)`:`Safety.(0-10)` + `Jobs.(0-10)`:`Civic.engagement.(0-10)` +
                  `Jobs.(0-10)`:`Accessiblity.to.services.(0-10)` + `Jobs.(0-10)`:`Housing.(0-10)` +
                  `Safety.(0-10)`:`Environment.(0-10)` +
                  `Safety.(0-10)`:`Civic.engagement.(0-10)` +
                  `Safety.(0-10)`:`Accessiblity.to.services.(0-10)` +
                  `Safety.(0-10)`:`Housing.(0-10)` +
                  `Accessiblity.to.services.(0-10)`:`Community.(0-10)`,
                  data=data_filtered %>% select(-Country, -Region, -Code))
summary(fit_model_3)

```

This model shows that “Education.(0-10)”, “Safety.(0-10)”,  “Civic.engagement.(0-10)”,  “Community.(0-10)”,  “Income.(0-10)”,  “Environment.(0-10)”,  “Safety.(0-10)”:”Accessiblity.to.services.(0-10)”,  “Education.(0-10)”:”Safety.(0-10)”,  “Education.(0-10)”:”Accessiblity.to.services.(0-10)”,  “Education.(0-10)”:”Housing.(0-10)”,  “Jobs.(0-10)”:”Housing.(0-10)”“Jobs.(0-10)”:”Income.(0-10)”,  “Jobs.(0-10)”:”Civic.engagement.(0-10)”,  “Safety.(0-10)”:”Environment.(0-10)”,  “Safety.(0-10)”:”Civic.engagement.(0-10)”,  “Safety.(0-10)”:”Housing.(0-10)”,  “Accessiblity.to.services.(0-10)”:”Community.(0-10)” shows significant contribution in life satisfaction scores.

Further, the model has a low MSE as compared to the simple model.

######################

Parameters from the three models-

### Adjusted R-square
```{r}
summary(fit_model_1)$adj.r.squared
summary(fit_model_2)$adj.r.squared
summary(fit_model_3)$adj.r.squared
```
### AIC
```{r}
AIC(fit_model_1)
AIC(fit_model_2)
AIC(fit_model_3)
```
### BIC
```{r}
BIC(fit_model_1)
BIC(fit_model_2)
BIC(fit_model_3)
```
### Residuals
```{r}
plot(fit_model_1$residuals)
plot(fit_model_2$residuals)
plot(fit_model_3$residuals)

```
### Normality using Q-Q Plots

```{r}
# Check for normality of residuals
# Model 1
hist(fit_model_1$residuals, breaks = 20)
qqnorm(fit_model_1$residuals)
qqline(fit_model_1$residuals)

# Model 2
hist(fit_model_2$residuals, breaks = 20)
qqnorm(fit_model_2$residuals)
qqline(fit_model_2$residuals)


# Model 3
hist(fit_model_2$residuals, breaks = 20)
qqnorm(fit_model_2$residuals)
qqline(fit_model_2$residuals)

```

###### ANOVA to check which one is better from Model 1 and Model 3

```{r}

anova(fit_model_1, fit_model_3)

```

Since the p-value is 2.2e-16 <<<< 0.05, we reject the Null Hypothesis and conclude that the Model 3 with Interaction Terms is the preferred model.


## Since the data has hierarchical strucutre, we fit the random intercept model.

###### Random Intercept Model

```{r}

fit_mem_ri <- lmer(
  `Life.satisfaction.(0-10)` ~ `Education.(0-10)` + `Jobs.(0-10)` + `Income.(0-10)` + 
                             `Safety.(0-10)` + `Health.(0-10)` + `Environment.(0-10)` + 
                             `Civic.engagement.(0-10)` + 
                            `Accessiblity.to.services.(0-10)` + 
                             `Housing.(0-10)` + `Community.(0-10)` + 
                             (1 | Country), 
  data = data_filtered %>% select(-Region, -Code)
)

# Summary of the model
summary(fit_mem_ri)

aic_model_4 <- AIC(fit_mem_ri)
bic_model_4<-  BIC(fit_mem_ri)


```

Significant Effects -> Education, Environment, Community (+ve), and Civic engagement (-ve).
Non-significant Effects -> Jobs, Income, Safety, Health, Housing, and Accessibility to services.

REML Criterion: 1131.8. Measure of model fit for restricted maximum likelihood (REML) estimation. Lower values of the REML criterion indicate a better-fitting model.



######################################################################### 
### Fitting the Random Slope and Random Intercept Model

###### Random Intercept Model with Random Slope for Education
```{r}
fit_mem_ri_slope_edu <- lmer(
  `Life.satisfaction.(0-10)` ~ `Education.(0-10)` + `Jobs.(0-10)` + `Income.(0-10)` + 
                             `Safety.(0-10)` + `Health.(0-10)` + `Environment.(0-10)` + 
                             `Civic.engagement.(0-10)` + `Accessiblity.to.services.(0-10)` + 
                             `Housing.(0-10)` + `Community.(0-10)` + 
                             (1 + `Education.(0-10)` | Country),  # Random intercept and slope for Education
  data = data_filtered %>% select(-Region, -Code)
)

summary(fit_mem_ri_slope_edu)

```
### Random Intercept Model with Random Slope for Community
```{r}
fit_mem_ri_slope_comm <- lmer(
  `Life.satisfaction.(0-10)` ~ `Education.(0-10)` + `Jobs.(0-10)` + `Income.(0-10)` + 
                             `Safety.(0-10)` + `Health.(0-10)` + `Environment.(0-10)` + 
                             `Civic.engagement.(0-10)` + `Accessiblity.to.services.(0-10)` + 
                             `Housing.(0-10)` + `Community.(0-10)` + 
                             (1 + `Community.(0-10)` | Country),  # Random intercept and slope for Education
  data = data_filtered %>% select(-Region, -Code)
)

summary(fit_mem_ri_slope_comm)

```

######################################################################### 

Plots for Random Intercept Model

```{r}
# plotting the predicted and the original values on separate regression plots.

# Create a dataframe with the original data and predicted values
data_with_pred <- cbind(data_filtered, pred = predict(fit_mem_ri))

ggplot(data_with_pred, aes(x = `Education.(0-10)`, y = `Life.satisfaction.(0-10)`, colour = Country)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = pred), method = "lm", se = FALSE, linewidth = 1) + 
  facet_wrap(~ Country) + 
  theme_minimal() + 
  labs(title = "Regression Fit Lines for All Countries", x = "Education (0-10)", y = "Life Satisfaction (0-10)")

```

Plots for Random Intercept Model with Random Slope for Education

```{r}
# plotting the predicted and the original values on separate regression plots.

# Create a dataframe with the original data and predicted values
data_with_pred_1 <- cbind(data_filtered, pred = predict(fit_mem_ri_slope_edu))

ggplot(data_with_pred, aes(x = `Education.(0-10)`, y = `Life.satisfaction.(0-10)`, colour = Country)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = pred), method = "lm", se = FALSE, linewidth = 1) + 
  facet_wrap(~ Country) + 
  theme_minimal() + 
  labs(title = "Regression Fit Lines for All Countries", x = "Education (0-10)", y = "Life Satisfaction (0-10)")
```


Plots for Random Intercept Model with Random Slope for Community

```{r}
# plotting the predicted and the original values on separate regression plots.

# Create a dataframe with the original data and predicted values
data_with_pred_1 <- cbind(data_filtered, pred = predict(fit_mem_ri_slope_comm))

ggplot(data_with_pred, aes(x = `Education.(0-10)`, y = `Life.satisfaction.(0-10)`, colour = Country)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = pred), method = "lm", se = FALSE, linewidth = 1) + 
  facet_wrap(~ Country) + 
  theme_minimal() + 
  labs(title = "Regression Fit Lines for All Countries", x = "Education (0-10)", y = "Life Satisfaction (0-10)")
```


###

```{r}
library(MuMIn)
r.squared_ri <- r.squaredGLMM(fit_mem_ri)

r.squared_ri_s_e <- r.squaredGLMM(fit_mem_ri_slope_edu)

r.squared_ri_s_c <- r.squaredGLMM(fit_mem_ri_slope_comm)

# Table for data

r_squared_table <- data.frame(
  Model = c("Random Intercept Model", 
            "Random Intercept with Random Slope for Education", 
            "Random Intercept with Random Slope for Community"),
  R_Squared_Marginal = c(r.squared_ri[1], r.squared_ri_s_e[1], r.squared_ri_s_c[1]),
  R_Squared_Conditional = c(r.squared_ri[2], r.squared_ri_s_e[2], r.squared_ri_s_c[2])
)

print(r_squared_table)


```

The table shows improvement in explanation of variables by the model with a random slope for Community. Further, it is noteworthy that the improvement over base models is not significant.


### Residual Analysis

```{r}

plot(fit_model_1$residuals)

```

```{r}
# Check for homoscedasticity (constant variance) by plotting residuals vs fitted values
plot(fitted(fit_model_1), residuals(fit_model_1))

```

```{r}

# Check for normality of residuals
hist(fit_model_1$residuals, breaks = 20)
qqnorm(fit_model_1$residuals)
qqline(fit_model_1$residuals)

```


```{r}

# Residuals for Model 4 (Mixed-Effects Model)
residuals_mem <- residuals(fit_mem_ri)

# Plot residuals
plot(residuals_mem)

# Check for homoscedasticity (plot residuals vs fitted values)
plot(fitted(fit_mem_ri), residuals_mem)

# Check for normality of residuals (Histogram and Q-Q plot)
hist(residuals_mem, breaks = 20)
qqnorm(residuals_mem)
qqline(residuals_mem)

```
```{r}
# Fitted values for Model 4 (Mixed-Effects Model)
fitted_values_mem <- fitted(fit_mem_ri)

# Plot residuals vs fitted values for homoscedasticity
plot(fitted_values_mem, residuals_mem)
```

### Creating Dot PLots

```{r}

library(lattice)
# Extract random effects
random_effects_mem <- ranef(fit_mem_ri)

# Plot random effects (use the random intercept and slope values)
dotplot(random_effects_mem)
```


```{r}

library(lattice)
# Extract random effects
random_effects_mem_ed <- ranef(fit_mem_ri_slope_edu)

# Plot random effects (use the random intercept and slope values)
dotplot(random_effects_mem_ed)

```


```{r}

library(lattice)
# Extract random effects
random_effects_mem_comm <- ranef(fit_mem_ri_slope_comm)

# Plot random effects (use the random intercept and slope values)
dotplot(random_effects_mem_comm)

```

AIC and BIC of the four models

```{r}

# Extract AIC and BIC for each model
aic_model_1 <- AIC(fit_model_1)
bic_model_1 <- BIC(fit_model_1)

aic_model_2 <- AIC(fit_model_2)
bic_model_2 <- BIC(fit_model_2)

aic_model_3 <- AIC(fit_model_3)
bic_model_3 <- BIC(fit_model_3)

aic_model_4 <- AIC(fit_mem_ri)
bic_model_4 <- BIC(fit_mem_ri)

aic_model_5 <- AIC(fit_mem_ri_slope_edu)
bic_model_5 <- BIC(fit_mem_ri_slope_edu)

aic_model_6 <- AIC(fit_mem_ri_slope_comm)
bic_model_6 <- BIC(fit_mem_ri_slope_comm)

# BIC Skipped, as we created complex models, and BIC always favors simple model


# Comparison Table
comparison_table <- data.frame(
  Model = c("Model 1", "Model 2", "Model 3", "Model 4 (Random Intercept)", "Model 5 (RI+S_Edu", "Model 6 (RI+S_Comm)"),
  AIC = c(aic_model_1, aic_model_2, aic_model_3, aic_model_4, aic_model_5, aic_model_6),
  BIC = c(bic_model_1, bic_model_2, bic_model_3, bic_model_4, bic_model_5, bic_model_6)
)

comparison_table

```
From this comparison, we see that the Interaction effect no group model gives a better fit for the data, instead of the complex models. This can be due to the fact that we lack enough data points in each category, thus making the complex models work poorly.









